---
title: "Homework 1"
author: "Helene, Rui, Hong"
date: "September 12, 2016"
output: 
  html_document:
    toc: yes
    toc_depth: 4
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=4, fig.height=3.5, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

### Question 1

### Question 2: The Radio Survey Dataset

#### a) Import and clean the data


 Read the data into R 
```{r, import, echo = F}
## Clear the R environment, and import necessary packages. 
rm(list=ls())
library(plyr)
library(ggplot2)
library(sqldf)
survey_raw <- read.csv("/Users/Rui/Dropbox/Upenn/Stat571/Data/Survey_results_final.csv")

options(scipen=1, digits=2)

```

Extract only the variables of interest.
```{r extract_variables}
cols_keep <- c("WorkTimeInSeconds", "Answer.Age", "Answer.Education", "Answer.Gender", "Answer.HouseHoldIncome", "Answer.Sirius.Radio", "Answer.Wharton.Radio")
survey <- survey_raw[cols_keep]
colnames(survey) <- c("worktime", "age", "edu", "gender", "income", "sirius", "wharton")
```



```{r cleaning }
## cleaning the gender variable
print("Answers given to Gender.")
levels(survey$gender)
survey <- survey[-c(which(survey$gender == "")),]
## cleaning the age variable
print("Answers given to Age.")
levels(survey$age)
survey <- survey[-which(survey$age == ""),]
survey <- survey[-which(survey$age == "female"),]
survey$age[which(survey$age == "223")] <- "23"
survey$age[which(survey$age == "27`")] <- "27"
survey$age[which(survey$age == "4")] <- "40"  # not sure if should delete this data point
survey$age[which(survey$age == "Eighteen (18)")] <- "18"
survey$age <- as.numeric(levels(survey$age))[survey$age]
## cleaning up the education variable
print("Ansers given to education")
unique(survey$edu)
survey <- survey[-which(survey$edu == "select one"),]
```

Clean the data

* Remove data points with gender ""

* Remove data points with age = "", "female"

* Replace mistyped age = "223", "4", "Eighteen (18)", "27`"

* Remove data points with education = "select one"


#### b) summarize all the variables collected 

**Gender **

```{r summarize_gender}
a <- summary(survey$gender)[2:3]
gender <- c("female", "male")
number <- c(a[1]/(a[1] + a[2]),a[2]/(a[1] + a[2]))
gender <- data.frame(gender, number)

p <- ggplot(gender, aes(x = 1, y = number, fill = gender)) + 
  geom_bar(stat="identity") +
  coord_polar(theta = 'y')
p <- p + 
  geom_bar(stat = "identity", color = 'black') +
  guides(fill=guide_legend(override.aes=list(colour=NA))) +
  theme(axis.ticks=element_blank(),  # the axis ticks
          axis.title=element_blank(),  # the axis labels
          axis.text.y=element_blank()) # the 0.75, 1.00, 1.25 labels.
print(p)
```

58% of the sample is male. 

**Work time **
```{r summarize_worktime}
summary(survey$worktime)
f1 <- ggplot(survey, aes(worktime, fill = gender)) + 
  geom_histogram(bins = 60) +
  xlab("Work Time (s)") + ylab("Count") + 
  ggtitle("Distribution of work time")
f1
```
**Age **
```{r summarize_age}
summary(survey$age)
sd(survey$age)
f2 <- ggplot(survey, aes(age, fill = gender)) + 
  geom_histogram(bins = 30) +
  xlab("Age") + ylab("Count") + 
  ggtitle("Distribution of participant age")
f2
```
**Education **
```{r summarize_education}
survey$edu <- revalue(survey$edu, c("Bachelor’s degree or other 4-year degree" = "Bachelor", 
                      "Graduate or professional degree" = "Graduate",
                      "High school graduate (or equivalent)" = "High school",
                      "Less than 12 years; no high school diploma" = "<12",
                      "Other" = "Other",
                      "Some college, no diploma; or Associate’s degree" = "Some college",
                      "select one" = "select one"   
                      ))
summary(survey$edu)
f <- ggplot(survey, aes(x = edu, fill = gender)) + geom_bar(position = "dodge")
print(f)
```

**Income **
```{r summarize_income}
survey$income_mean <- NA
survey$income_mean[which(survey$income == "$15,000 - $30,000")] <- mean(c(15000, 30000))
survey$income_mean[which(survey$income == "$30,000 - $50,000")] <- mean(c(30000, 50000))
survey$income_mean[which(survey$income == "$50,000 - $75,000")] <- mean(c(50000, 75000))
survey$income_mean[which(survey$income == "$75,000 - $150,000")] <- mean(c(75000, 150000))
survey$income_mean[which(survey$income == "Above $150,000")] <- (150000)
survey$income_mean[which(survey$income == "Less than $15,000")] <- (15000)
survey$income_mean <- as.numeric(survey$income_mean) 
table(survey$income_mean)

ggplot(survey,aes(income_mean, fill = gender)) + 
  geom_bar(position = "dodge")
```


### b) Sample properties
#### b.1) Sample vs. U.S. population
Does this sample appear to be a random sample from the general population of the U. S.? IF so we can get our estimate easily. (Utilize your control variables, compare them to sources such as the CIA world factbook).
```{r sample_US}
n <- length(survey$age) # total sample size
p1 <- length(which(survey$age <25))
p2 <- length(which(survey$age < 55)) - p1
p3 <- length(which(survey$age < 65)) - p1 - p2
p4 <- length(which(survey$age > 64))

df1 <- survey[which(survey$age < 25),]
d1 <- summary(df1$gender)
df2 <- survey[setdiff(which(survey$age < 55), which(survey$age < 25)),]
d2 <- summary(df2$gender)
df3 <- survey[setdiff(which(survey$age < 65), which(survey$age < 55)),]
d3 <- summary(df3$gender)
df4 <- survey[which(survey$age > 64),]
d4 <- summary(df4$gender)

```


Age group | CIA age structure (\%) | Sample age structure (\%) | CIA sex ratio (males/female) | Sample sex ratio (males/female) 
------------- | -------------  | -------------  | -------------  | ------------- 
0 - 14 years      | 18.99    | 0           | 1.04            | NA     
15 - 24 years     | 13.64   |    `r 100 * p1/n`     | 1.05      | `r  d1[3]/d1[2]`                    
25 - 54 years     | 39.76   |  `r 100 *p2/n`    | 1          |   `r d2[3]/d2[2]`
55 - 64 years     | 12.73     | `r 100 *p3/n`   | 0.93     |     `r d3[3]/d3[2]`                          
65 years and over | 14.88  |  `r 100 *p4/n` | 0.79            |   `r d4[3]/d4[2]`                        

#### b 2) Sample vs. mTurk population
Does this appear to be a random sample from mTurk population?

We downloaded demographic data of 5000 most recent mTurk surveys from https://mturk-surveys.appspot.com/, and compared this demographic information with our sample. 
```{r sample_mturk}
mturk <- read.csv("/Users/Rui/Dropbox/Upenn/Stat571/Homework/HW1/mturk_surveys.csv")

n_mturk <- length(mturk$Age)
mf1 <- mturk[which(mturk$Age < 25),]
m1 <- summary(mf1$gender)
mf2 <- mturk[setdiff(which(mturk$Age < 55),which(mturk$Age < 25)),]
m2 <- summary(mf2$gender)
mf3 <- mturk[setdiff(which(mturk$Age < 65),which(mturk$Age < 55)),]
m3 <- summary(mf3$gender)
mf4 <- mturk[which(mturk$Age > 64),]
m4 <- summary(mf4$gender)

```

Age group | MTurk age structure (\%) | Sample age structure (\%) | MTurk sex ratio (males/female) | Sample sex ratio (males/female) 
------------- | -------------  | -------------  | -------------  | ------------- 
16 - 24 years     |  `r 100*length(mf1$Age)/n_mturk`  |   `r 100 * p1/n`     | `r  m1[2]/m1[1]` | `r  d1[3]/d1[2]`                    
25 - 54 years     |  `r 100*length(mf2$Age)/n_mturk`  |   `r 100 * p2/n`   | `r  m2[2]/m2[1]`    |   `r d2[3]/d2[2]`
55 - 64 years     | `r 100*length(mf3$Age)/n_mturk`    |  `r 100 * p3/n` |`r  m3[2]/m3[1]`    |     `r d3[3]/d3[2]`                          
65 years and over |  `r 100*length(mf4$Age)/n_mturk` |  `r 100 * p4/n` | `r  m4[2]/m4[1]`   |   `r d4[3]/d4[2]`  


This table suggests that our sample is a fairly good representation of Amazon Mechanical Turk U.S. population.

#### b  3) Estimating U.S. Wharton radio listeners.
Assume that the proportion of Wharton listeners vs. that of Sirius listeners remains the same in the general population as it is in the MTURK population. Use the data to provide an estimate of the number of Wharton listeners in the U. S. In order to make this estimate do you need to break down the proportion of Wharton to Sirius by age? Provide some graphical or numerical evidence to support your reasoning.

We break down the proportion of Wharton to Sirius listeners by age.  **Here we are assuming that sample represents MTurk population **

Age group |  CIA age structure (\%) | MTurk age structure (\%) |  Wharton to Sirius listeners (\%)
------------- | -------------  | -------------  |  ------------- 
16 - 24 years    | 13.64  |  `r 100*length(mf1$Age)/n_mturk`  | `r summary(df1$wharton)[3]/summary(df1$sirius)[3]`                   
25 - 54 years    | 39.76  |  `r 100*length(mf2$Age)/n_mturk`  | `r summary(df2$wharton)[3]/summary(df2$sirius)[3]`
55 - 64 years    | 12.73 | `r 100*length(mf3$Age)/n_mturk`   |  `r summary(df3$wharton)[3]/summary(df3$sirius)[3]`
65 years and over | 14.88 |  `r 100*length(mf4$Age)/n_mturk`  |  `r summary(df4$wharton)[3]/summary(df4$sirius)[3]`

One approach is to estimate the proportion of the Wharton listeners to that of the Sirius listeners, p, so that we will come up with an audience size estimate of approximately 51.6 million times p.

Hence, we estimate the proportion of the Wharton listeners to that of the Sirius listeners in the U.S 
$$ p = 13.64\% \times 0.06 + 39.76\% \times 0.05 + 12.73\% \times 0.03 + 14.88\% \times 0.14 $$
p = `r 0.1364*0.06 + 0.3976*0.05 + 0.1273*0.03 + 0.1488*0.14`

Number of wharton listeners =  $p \times 51.6 =  2.6 million$ 

### Question 3 

### Question 4 : Major League Baseball

####i) Summarize the data by providing suitable summary statistics and graphs. Write a brief report for your findings.

#####Load and examine dataset:
paydata <- read.csv("MLPayData_Total.csv")
names(paydata)

#####select the variables we are interested in: team names, team total payroll from 1998 to 2014 (in billions), fraction of games won from 1998 to 2014, and payroll (in millions) and fraction of games won for the years we want to examine:  
paydata1 <- paydata[,c(3,1:2,4,54,12,46,20,38)]
head(paydata1)

#####rename the variables to more intelligible names: 
names(paydata1) <- c("team","tot.payroll","avgwin","payroll.1998","pctwin.1998","payroll.2006","pctwin.2006","payroll.2014","pctwin.2014")
head(paydata1)

#####create a vector of the mean of each variable, beside team 
tot.mean <- apply(X = paydata1[,-1], MARGIN = 2, FUN = mean)

#####create tot.mean row with the mean for all variables............. work in progress.................. it's not working!!!!!!! 
team <- "tot.mean"
tot.mean <- rbind(team, tot.mean)

paydata2 <- rbind(paydata1, tot.mean)

tot.mean <- c("tot.mean",tot.mean)
paydata1[,1]

#####display the observations for the teams we are interested in (Boston Red Sox & Oakland A's), plus the first three teams in the dataset.  
paydata1[c(4,20,1,2,3),]

#####boxplots for tot.payroll and avgwin 
boxplot(paydata$payroll)
######WORK IN PROGRESS......trying to put a horizontal line for A's and Red Sox 
  abline(
    h = 0.5487172,
    col = "lightblue",
    lwd = 2 
  )
######END OF WORK IN PROGRESS....... 
boxplot(paydata$avgwin)

#####scatterplot of average by total payroll from 1998 to 2014 
plot(paydata$payroll, paydata$avgwin, 
       pch  = 16, 
       cex  = 1.2,
       col  = "blue",
       xlab = "Payroll", 
       ylab = "Win Percentage",
       main = "MLB Teams's Overall Win Percentage vs. Payroll")

######WORK IN PROGRESS: identify the Oakland A's and Boston Red Sox in the scatterplot. 
identify(paydata$payroll, paydata$avgwin, labels=paydata$team, plot=TRUE) 

####ii) For a given year, is payroll a significant variable to predict the winning percentage of that year? Choose a year and run a regression to examine this. You may try this for a few different years. Explain your results.

  myfit0 <- lm(avgwin~tot.payroll, data=paydata1)
  summary(myfit0)
  
  myfit1998 <- lm(pctwin.1998~payroll.1998, data=paydata1)  
  summary(myfit1998)
  
  myfit2006 <- lm(pctwin.2006~payroll.2006, data=paydata1)  
  summary(myfit2006)
  
  myfit2014 <- lm(pctwin.2014~payroll.2014, data=paydata1)  
  summary(myfit2014)
  
  par(mfrow = c(2,2))
  
    plot(paydata1$tot.payroll, paydata$avgwin, 
       pch  = 16, 
       cex  = 1.2,
       col  = "blue",
       xlab = "Payroll", 
       ylab = "Win Percentage",
       xlim = c(0, 2.5),
       ylim = c(0.4, 0.7),
       main = "MLB Teams's Overall Win Percentage vs. Payroll")
    abline(myfit0, col="red", lwd=4)
    
    plot(paydata1$payroll.1998, paydata1$pctwin.1998, 
       pch  = 16, 
       cex  = 1.2,
       col  = "blue",
       xlab = "Payroll", 
       ylab = "Win Percentage",
       xlim = c(0, 250),
       ylim = c(0.4, 0.7),
       main = "MLB Teams's Win Percentage vs. Payroll in 1998")
    abline(myfit1998, col="red", lwd=4) 

    plot(paydata1$payroll.2006, paydata1$pctwin.2006, 
       pch  = 16, 
       cex  = 1.2,
       col  = "blue",
       xlab = "Payroll", 
       ylab = "Win Percentage",
       xlim = c(0, 250),
       ylim = c(0.4, 0.7),
       main = "MLB Teams's Win Percentage vs. Payroll in 2006")
    abline(myfit2006, col="red", lwd=4) 
  
    plot(paydata1$payroll.2014, paydata1$pctwin.2014, 
       pch  = 16, 
       cex  = 1.2,
       col  = "blue",
       xlab = "Payroll", 
       ylab = "Win Percentage",
       xlim = c(0, 250),
       ylim = c(0.4, 0.7),
       main = "MLB Teams's Win Percentage vs. Payroll in 2014")
    abline(myfit2014, col="red", lwd=4) 
  
####iii) Now use the aggregated information, and use regression to analyze the relation between total payroll and overall winning percentage. Run appropriate model(s) to answer the following questions.
#####a) In this analysis do the Boston Red Sox perform reasonably well given their total payroll? [Use a 95% prediction interval.]

  attach(paydata1)
  
  new <- data.frame(tot.payroll=c(1.972359)) #Boston Red Sox's total payroll 
  CImean <- predict(myfit0, new, interval="prediction", se.fit=TRUE)  
  CImean

######Answer: The Boston Red Sox, with an average of 55% of games won between 1998 and 2014, falls within the confidence interval in terms of their performance given their payroll.

#####b) In view of their winning percentage, how much payroll should the Oakland A’s have spent? [Use a 95% prediction interval.]

  myfit1 <- lm(tot.payroll~avgwin, data=paydata1)
  summary(myfit1)

  new2 <- data.frame(avgwin=c(0.5445067)) #Oakland A's average of games won between 1998 and 2014
    CImean <- predict(myfit1, new2, interval="prediction", se.fit=TRUE)  
    CImean

######Answer: Using the regression model in myfit1, we would have predicted that a team winning 54% of all games between 1998 and 2014 should have a total payroll over that time between 0.95 and 2.27 billion dollars. Given this result, we can say that the Oakland A's performed beyond expectation for their budget. 

#####iv) What is the best model you could build on the basis of this data to capture the important factors affecting winning percentage in 2014 and predict that percentage for 2015? 
#####Describe the process of your analysis. What criterion are you using? Report your findings.

  myfit2014 <- lm(pctwin.2014~payroll.2014, data=paydata1)  
  summary(myfit2014)

  attach(paydata1)
  
  new3 <- data.frame(payroll.2014=c(162.8174)) #Boston Red Sox's payroll for 2014
  CImean <- predict(myfit2014, new3, interval="prediction", se.fit=TRUE) #Prediction of Boston Red Sox's performance in 2015 based on the myfit2014 model, and its payroll in 2014 
  CImean
  
######I really don't know whether this is the right way. Please let me know if you have a better idea! 


### Question 5 